---
title: "PCA"
author: "Ruiqi (Rickey) Huang"
date: "2/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
#install.packages("rlang")
library(rlang)
```

```{r,echo=FALSE}
library(ggplot2)
#install.packages("ggfortify")
#install.packages("factoextra")
library(ggfortify)
library(factoextra)
```

## Read and summarize the data

```{r}
TimesRanking <- read.csv("TimesCountryAvg.csv")
#dim(TimesRanking)
head(TimesRanking)
```

## Create a unified scale for all possible variables

### Create a covariance matrix S for our data

  After calculating all the covariances, we could find the *total variance* by sum all variance we have (we calculate the $\sum_{j=1}^{k}S_{jj}$, and this is actually the sum of eigenvalues of S)
  
```{r}
S <- cov(TimesRanking[, 6:13])
S
sum(diag(S))
```

### find eigenvalues and eigenvectors of S

```{r}
s.eigen <- eigen(S)
s.eigen
```

The eigenvectors represent the principal components of S. The eigenvalues of S are used to find the proportion of the total variance explained by the components.

```{r}
for (s in s.eigen$values) {
  print(s / sum(s.eigen$values))
}
```

The first two principal components account for 88.3% of the total variance. 

(*TODO: Here we need to know what level of total variance we want our model to explain*)

A scree graph of the eigenvalues can be plotted to visualize the proportion of variance explained by each subsequential eigenvalue.

```{r}
plot(s.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph')
lines(s.eigen$values)
```

```{r}
s.eigen$vectors
```

### use the eigenvectors to find out the results

The elements of the eigenvectors of S are the 'coefficients' or 'loadings' of the principal components.

here if we decide we are going to use the first two components as the principal components we can write out the formula for these two principal compnents from the eigenvectors result above:

$1^{st}$ principal component 

$y_{1} = 0.03118182 * StudentToStaffAvg + 0.0361542981 *  intlStudentAvg +0.9784538847 * femalePersentageAvg + 0.144736992 * malepercentageAvg - 0.139241131*teachingAvg- 0.0016289660*reseachAvg + 0.0024238021*citation + 0.001771678*industryIncomeAvg$

$y_{1} = 0.002569285 * StudentToStaffAvg -0.0014031869 *  intlStudentAvg -0.0010704334* femalePersentageAvg + 0.0036249392 * malepercentageAvg + 0.003512949*teachingAvg- 0.9623460710*reseachAvg - 0.2691096581*citation + 0.03787921e*industryIncomeAvg$

### Analyze the principal components

From the result above the $y_{1}$ combines variables StudentToStaffAvg, intlStudent, femalePercentageAvg, and teaching, while the component $y_{2}$ measures the change in the variables like reserchAvg, citation, and industryIncomAvg.

In this case, we might name our component $y_{1}$ as  Demographical factor and name the second component $y_{2}$ as Educational Outcome factor.

## Another method to do PCA

```{r}
TimesRanking.pca <- prcomp(TimesRanking[,6:13], scale = TRUE)
TimesRanking.pca
summary(TimesRanking.pca)
```


## Plot the principal components

```{r}
TR.PCA.plot <- autoplot(TimesRanking.pca, data = TimesRanking, colour = 'scoreAvg')
TR.PCA.plot
TR.PCA.plot2 <- autoplot(TimesRanking.pca, data = TimesRanking, colour = 'Health')
TR.PCA.plot2
```

*TODO: Here we still need to find some possible categorical or numerical response variable (output) to measure the level of health (may be some indirectly related variables), in order to chack whether our PC1 & 2 like above.*

In the plots above, I used the scoreAvg (Times ranking score) to show the cluster in the first plot, which shows that a higher score means a combination of a high PC1 and a high PC2. For the seconf plot, I manually labelled countries with score from 40 to the max as healthy, countries with score from 20 to 40 as neet to improve, and countries from min to 20 as unhealthy. The plot shows that the countries with "unhealthy" education system have a low PC1 and a median PC2, and the countries "need improvement" in education have a median PC1 and a low or median PC2, while the countries with a "healthy" education sytem have a high PC1 and a median or high PC2. This way of evaluating Health variable may not be practical, some more reasonable response variable need to be chosen.


## Visualizations

```{r}
TR.Individuals <- fviz_pca_ind(TimesRanking.pca,
                               col.ind = "cos2", #color by the quality of representation
                               gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), #set up the colors
                               reprel = TRUE #avoid test overlapping
                               )
TR.Individuals
```

